---
title: "Question 2 - Association Mining"
author: "Evan Canfield"
date: "7/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Library Calls

The following libraries are used to develop the data frames for Question 1 of the Group Project

```{r library, warning=FALSE, message=FALSE}
if(!require(pacman)){install.packages("pacman")}
library(pacman)

p_load(  
  arules,
  arulesViz,
  caTools,
  Hmisc, 
  kableExtra,
  plotly,
  skimr,
  tidyverse
)
```

# Data Import
```{r cache=TRUE}
#Processed QVC Data Frame
qvc_data <- readRDS(file = "./data/output/Group Project Dataset/qvc_data_compiled_new_variables.rds")
```

## Check Data Import
### QVC Data
```{r}
glimpse(qvc_data)
```

# Subset Data For Quicker Process Times
```{r}
# set.seed(101)
# 
# split_ratio <- 0.30
# 
# sample <- sample.split(Y = qvc_data$Sales_Order_Nbr
#                        ,SplitRatio =  split_ratio)
# 
# qvc_data_sample = subset(qvc_data, sample == TRUE)
```

```{r}
# qvc_data <- qvc_data_sample
```



# Select Essential Variables
```{r}
qvc_data_am <- qvc_data %>% 
  filter(Merchandise_Dept_Desc == "Health/Beauty") %>% 
  select(Sales_Order_Nbr,
         Product_Id)

glimpse(qvc_data_am)
```


# For Checking - Item Counts
```{r}
product_counts <- qvc_data_am %>% 
  group_by(Product_Id) %>% 
  summarise(total = n()) %>% 
  arrange(-total) %>% 
  filter(total > 1)

head(product_counts)
```


# Export / Import Transactions
```{r}
write.csv(x = qvc_data_am
        , file = "./data/output/q2_association_mining/qvc_data_sales-and-prod.csv"
        , row.names = FALSE)

qvc_sales_trans <- read.transactions(file = "./data/output/q2_association_mining/qvc_data_sales-and-prod.csv"
                  , format = "single"
                  , cols = c(1,2)
                  , sep = ","
                  , rm.duplicates = TRUE
                  , skip = 1)
```

# Inspect
```{r}
summary(qvc_sales_trans)

qvc_sales_trans@data

str(qvc_sales_trans)

qvc_sales_trans@itemInfo$labels[1:20]

qvc_sales_trans@data@p
```

# Apriori Analysis
## Frequent Sets
```{r}
frqsets.qvc <- apriori(data = qvc_sales_trans
                       , parameter=list(minlen=2
                                        , supp=1e-5
                                        , conf=0.5
                                        , target="frequent itemsets")
                       , control = list(verbose = FALSE))

frqsets.qvc

frqsets.qvc_df <- inspect(frqsets.qvc)

frqsets.qvc_df

 frqsets.qvc_df %>% 
  arrange(-count) %>% 
  head(10) %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")
                , full_width = FALSE)

```

## Rules
```{r}
rules.qvc <- apriori(data = qvc_sales_trans
                     , control = list(verbose=TRUE)
                     , parameter = list(minlen=2
                                        , supp = 1e-6
                                        , conf=0.5)) 

summary(rules.qvc)

inspect(rules.qvc)
```

# Redundant Rules
```{r}
redundant <- which (colSums(is.subset(rules.qvc, rules.qvc)) > 1)
```

## Prune Redundant Rules
```{r}
rules.qvc.pruned <- rules.qvc[-redundant]

rules.qvc.pruned
```

```{r}
subrules2 <- head(sort(rules.qvc.pruned, by="lift"), 10)

plot(subrules2, method="graph")
```

# Create Rule Data Frame
```{r}
rules.qvc_df <- data.frame(
  lhs = labels(lhs(rules.qvc.pruned)),
  rhs = labels(rhs(rules.qvc.pruned)),
  rules.qvc.pruned@quality
)

rules.qvc_df_10 <- rules.qvc_df %>% 
  arrange(-lift) %>% 
  mutate(confidence = round(confidence,3),
         lift = round(lift,0)) %>% 
  head(10)

rules.qvc_df_10 %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")
                , full_width = FALSE)

```