---
title: "Data Processing"
author: "Evan Canfield"
date: "7/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Library Calls

```{r library, warning=FALSE, message=FALSE}
if(!require(pacman)){install.packages("pacman")}
library(pacman)

p_load(tidyverse, caTools, Hmisc, rowr, readxl, skimr, stringr, zipcode, maps, tidycensus, lubridate)
```

## Data Import

The following files were provided for analysis of QVC fulfillment for the DSBA 6211 Group Project:

* QVC Data 1
* QVC Data 2
* QVC Data 3
* QVCdist_ctr
* QVCorderstatustype
* QVC Data Dictionary

These files, with the exception of "QVC Data Dictionary", will be uploaded for processing .

An error was noticed in the data during inspectiion in Excel. The rows listed below had blank value inserted at Size_DESC and the remaining data shifted over one column. Therefore, this shifted data was readjusted manually before upload.

* QVC Data 1: 220464 (#Sales_Order_Nbr: 649344509406)
* QVC Data 1: 259651 (#Sales_Order_Nbr: 449320937874)
* QVC Data 2: 111684 (#Sales_Order_Nbr: 649288839226)
* QVC Data 2: 462659 (#Sales_Order_Nbr: 849460177698)
* QVC Data 3: 271093 (#Sales_Order_Nbr: 149349466481)
* QVC Data 3: 303069 (#Sales_Order_Nbr: 649356909726)

```{r data import, warning=FALSE, cache.lazy=TRUE}
qvc_data_1 <- read_xlsx("./data/QVC Data 1.xlsx", na = c("NA", ""))

#qvc_data_2 <- read_xlsx("./data/QVC Data 2.xlsx", na = c("NA", ""))

#qvc_data_3 <- read_xlsx("./data/QVC Data 3.xlsx", na = c("NA", ""))

qvc_distctr <- read.csv("./data/QVCdist_ctr.csv")

qvc_orderstatustype <- read_xlsx("./data/QVCorderstatustype.xlsx", na = c("NA", "")) 
```

## Join the QVC Data Files

With all three QVC Data files available they can now be combined into a single dataframe.
```{r bind qvc data files}
# qvc_data <- qvc_data_1 %>% 
#   bind_rows(qvc_data_2) %>% #Bind QVC Data 2 to QVC Data 1
#   bind_rows(qvc_data_3)     #Bind QVC Data 3  

qvc_data <- qvc_data_1
```

## Subset Data for Faster Processing During Investigation
```{r}
set.seed(101)

split_ratio <- 0.20

sample <- sample.split(qvc_data$Sales_Order_Line_Nbr, split_ratio)

qvc_data = subset(qvc_data, sample == TRUE)

```

## Data Inspection

#### Dataframe Structure
```{r summary}
glimpse(qvc_data)
```

#### Data Summary
```{r describe}
#describe(qvc_data)
```

```{r skim dataset}
skim(qvc_data)
```

```{r}
summary(qvc_data)
```

## Rename Sales Order Number
```{r}
qvc_data <- qvc_data %>% 
  rename(Sales_Order_Nbr = '#Sales_Order_Nbr')
```


## Data Investigation
### Pair 2 Asssigned Variables
#### Sales_Order_Line_Nbr
Data Type Check / Conversion

Stats
```{r}
qvc_data$Sales_Order_Line_Nbr <- as.factor(qvc_data$Sales_Order_Line_Nbr)

qvc_data %>% 
  select(Sales_Order_Line_Nbr) %>% 
  skim()

qvc_data %>% 
  select(Sales_Order_Line_Nbr) %>% 
  summary()

qvc_data %>% 
  group_by(Sales_Order_Line_Nbr) %>%
  tally() %>% 
  mutate(Percent = round(n / nrow(qvc_data)*100, 3)) %>% 
  arrange(desc(n)) %>% 
  head(n = 10)
```

#### Party_Id
Data Type Check / Conversion
```{r}
qvc_data$Party_Id <- as.character(qvc_data$Party_Id)

qvc_data %>% 
  select(Party_Id) %>% 
  skim()

qvc_data %>% 
  select(Party_Id) %>% 
  summary()

qvc_data %>% 
  group_by(Party_Id) %>%
  tally() %>% 
  mutate(Percent = round(n / nrow(qvc_data)*100, 3)) %>% 
  arrange(desc(n)) %>% 
  head(n = 10)
```




#### Total_Line_Amt
Data Type Check / Conversion
```{r}
qvc_data %>% 
  select(Total_Line_Amt) %>% 
  skim()
```

Log Transformation Check
```{r}
#Log Transformation
qvc_data %>% 
  select(Total_Line_Amt) %>% 
  mutate(Total_Line_Amt_log = log(Total_Line_Amt + 1)) %>% 
  gather() %>% 
  ggplot(mapping = aes(x = value, group = key)) +
    geom_histogram(bins = 50) + 
    facet_wrap(~ key, scales = "free")
```

#### Line_Status_Cd
Data Type Check / Conversion
```{r}
qvc_data$Line_Status_Cd <- as.character(qvc_data$Line_Status_Cd)

qvc_data %>% 
  select(Line_Status_Cd) %>% 
  skim()

qvc_data %>% 
  select(Line_Status_Cd) %>% 
  summary()

qvc_data %>% 
  group_by(Line_Status_Cd) %>%
  tally() %>% 
  mutate(Percent = round(n / nrow(qvc_data)*100, 3)) %>% 
  arrange(desc(n)) %>% 
  head(n = 10)
```

#### Line_Status_Dt
Data Type Check / Conversion
```{r}
qvc_data %>% 
  select(Line_Status_Dt) %>% 
  skim()

qvc_data %>% 
  select(Line_Status_Dt) %>% 
  summary()
```

#### Sku_Id
Data Type Check / Conversion
```{r}
qvc_data %>% 
  select(Sku_Id) %>% 
  skim()

qvc_data$Sku_Id <- as.character(qvc_data$Sku_Id)
```

```{r}
qvc_data %>% 
  select(Sku_Id) %>% 
  skim()

qvc_data %>% 
  select(Sku_Id) %>% 
  summary()

qvc_data %>% 
  group_by(Sku_Id) %>%
  tally() %>% 
  mutate(Percent = round(n / nrow(qvc_data)*100, 3)) %>% 
  arrange(desc(n)) %>% 
  head(n = 10)
```

#### Shipped_Dt
Data Type Check / Conversion
```{r}
qvc_data %>% 
  select(Shipped_Dt) %>% 
  skim()

qvc_data %>% 
  select(Shipped_Dt) %>% 
  summary()


qvc_data %>% 
  filter(is.na(Shipped_Dt)) %>% 
  nrow() / nrow(qvc_data) * 100
```

Percentages
```{r}
#Percentage NA
qvc_data %>% 
  filter(is.na(Shipped_Dt)) %>% 
  nrow() / nrow(qvc_data) * 100

#Percentage 2017
qvc_data %>% 
  filter((Shipped_Dt < as.Date("2017-01-01"))) %>% 
  nrow() / nrow(qvc_data) * 100
```

Histogram
```{r}
qvc_data %>% 
  select(Shipped_Dt) %>% 
  ggplot(mapping = aes(x = Shipped_Dt))+
  geom_histogram()

mean.POSIXlt(qvc_data$Shipped_Dt)

mean.Date()

```

Mean Date
```{r}
qvc_data_Shipped_DT <- qvc_data %>% 
  select(Shipped_Dt) %>% 
  drop_na()

mean.Date(as.list(qvc_data_Shipped_DT))

mean.POSIXct(qvc_data_Shipped_DT)
```


#### Cancelled_Qty
Data Type Check / Conversion
```{r}
qvc_data %>% 
  select(Cancelled_Qty) %>% 
  skim()

qvc_data %>% 
  select(Cancelled_Qty) %>% 
  summary()

qvc_data %>% 
  select(Cancelled_Qty) %>% 
  describe()
```

#### Merchandise_Div_Desc
Data Type Check / Conversion
```{r}
qvc_data %>% 
  select(Merchandise_Div_Desc) %>% 
  skim()

qvc_data$Merchandise_Div_Desc <- as.factor(qvc_data$Merchandise_Div_Desc)
```

Stats
```{r}
qvc_data %>% 
  select(Merchandise_Div_Desc) %>% 
  skim()

qvc_data %>% 
  select(Merchandise_Div_Desc) %>% 
  summary()

qvc_data %>% 
  group_by(Merchandise_Div_Desc) %>%
  count() %>% 
  mutate(Percent = round(n / nrow(qvc_data)*100, 1)) %>% 
  arrange(desc(n))
```

#### Merchandise_Dept_Desc
Data Type Check / Conversion
```{r}
qvc_data %>% 
  select(Merchandise_Dept_Desc) %>% 
  skim()

qvc_data$Merchandise_Dept_Desc <- as.factor(qvc_data$Merchandise_Dept_Desc)
```

Stats
```{r}
qvc_data %>% 
  select(Merchandise_Dept_Desc) %>% 
  skim()

qvc_data %>% 
  select(Merchandise_Dept_Desc) %>% 
  summary()

qvc_data %>% 
  group_by(Merchandise_Dept_Desc) %>%
  count() %>% 
  mutate(Percent = round(n / nrow(qvc_data)*100, 1)) %>% 
  arrange(desc(n))
```

#### Merchandise - Compare
```{r}
qvc_data_merch_div <- qvc_data %>% 
  group_by(Merchandise_Div_Desc) %>%
  count() %>% 
  mutate(Percent = round(n / nrow(qvc_data)*100, 1)) %>% 
  arrange(desc(n))

qvc_data_merch_dept <- qvc_data %>% 
  group_by(Merchandise_Dept_Desc) %>%
  count() %>% 
  mutate(Percent = round(n / nrow(qvc_data)*100, 1)) %>% 
  arrange(desc(n))

cbind.fill(qvc_data_merch_div, qvc_data_merch_dept)
```

### Shipment_Status_Dt
Data Type Check / Conversion
```{r}
qvc_data %>% 
  select(Shipment_Status_Dt) %>% 
  skim()

qvc_data %>% 
  select(Shipment_Status_Dt) %>% 
  summary()
```


### Rescheduled_Delivery_Dt
Data Type Check / Conversion
```{r}
qvc_data %>% 
  select(Rescheduled_Delivery_Dt) %>% 
  skim()

qvc_data %>% 
  select(Rescheduled_Delivery_Dt) %>% 
  summary

#Percentage NA
qvc_data %>% 
  filter(is.na(Rescheduled_Delivery_Dt)) %>% 
  nrow() / nrow(qvc_data) * 100
```

#### Actual_Total_Package_Qty
Data Type Check / Conversion
```{r}
qvc_data %>% 
  select(Actual_Total_Package_Qty) %>% 
  skim()

qvc_data %>% 
  select(Actual_Total_Package_Qty) %>% 
  summary()

qvc_data %>% 
  select(Actual_Total_Package_Qty) %>% 
  describe()
```

Log Transformation Check
```{r}
#Log Transformation
qvc_data %>% 
  select(Actual_Total_Package_Qty) %>% 
  mutate(Actual_Total_Package_Qty_log = log(Actual_Total_Package_Qty + 1)) %>% 
  gather() %>% 
  ggplot(mapping = aes(x = value, group = key)) +
    geom_histogram(bins = 10) + 
    facet_wrap(~ key, scales = "free")
```

#### SHIP_TO_STATE
Data Type Check / Conversion
```{r}
qvc_data %>% 
  select(SHIP_TO_STATE) %>% 
  skim()

qvc_data$Merchandise_Div_Desc <- as.factor(qvc_data$Merchandise_Div_Desc)
```

Stats
```{r}
qvc_data %>% 
  select(SHIP_TO_STATE) %>% 
  skim()

qvc_data %>% 
  select(SHIP_TO_STATE) %>% 
  summary()

qvc_data %>% 
  select(SHIP_TO_STATE) %>% 
  describe()

qvc_data %>% 
  group_by(SHIP_TO_STATE) %>%
  count() %>% 
  mutate(Percent = round(n / nrow(qvc_data)*100, 1)) %>% 
  arrange(desc(n))
```

#### Zip Codes
```{r}
qvc_data %>% 
  select(SHIP_TO_ZIP) %>% 
  skim()

qvc_data_zip <- qvc_data

# Replace Dropped Leading Zeros
qvc_data_zip$SHIP_TO_ZIP <- str_pad(string = qvc_data_zip$SHIP_TO_ZIP, 
                                width = 5, 
                                side = "left", 
                                pad = "0")
# Verify Change
describe(qvc_data_clean_1.1$SHIP_TO_ZIP)

```
### Delivery_Confirmation_Dt
```{r}
qvc_data %>% 
  select(Delivery_Confirmation_Dt) %>% 
  skim()

#Percent NA
qvc_data %>% 
  filter(is.na(Delivery_Confirmation_Dt)) %>% 
  nrow() / nrow(qvc_data) * 100

#Missing Delivery Date
qvc_data_missing_delivery_dt <- qvc_data %>% 
  filter(is.na(Delivery_Confirmation_Dt)) %>% 
  mutate(Assigned_Dc_Id = as.character(Assigned_Dc_Id))

#Not missing delivery date
qvc_data_NOT_missing_delivery_dt <- qvc_data %>% 
  filter(!is.na(Delivery_Confirmation_Dt)) %>% 
  mutate(Assigned_Dc_Id = as.character(Assigned_Dc_Id))

skim(qvc_data_missing_delivery_dt)

#What are the order status for missing delivery dates
qvc_data_missing_delivery_dt %>% 
  group_by(Line_Status_Cd) %>% 
  count() 

qvc_data_missing_delivery_dt
```

Missing Delivery Date Visuals
```{r}
#What are the order status for missing delivery dates
qvc_data_missing_delivery_dt %>% 
  group_by( Line_Status_Cd) %>% 
  count() %>% 
  arrange(desc(n)) %>%
  mutate(n_percent = n/nrow(qvc_data_missing_delivery_dt)*100) %>% 
  ggplot(mapping = aes(x = Line_Status_Cd, y = n_percent)) +
  geom_col()

#What are the order status for missing delivery dates
qvc_data_NOT_missing_delivery_dt %>% 
  group_by( Line_Status_Cd) %>% 
  count() %>% 
  arrange(desc(n)) %>% 
  mutate(n_percent = n/nrow(qvc_data_NOT_missing_delivery_dt)*100) %>% 
  ggplot(mapping = aes(x = Line_Status_Cd, y = n_percent)) +
  geom_col()
```

#### Package_Cnt
```{r}
qvc_data %>% 
  select(Package_Cnt) %>% 
  skim()

qvc_data %>% 
  select(Package_Cnt) %>% 
  summary()

qvc_data %>% 
  select(Package_Cnt) %>% 
  describe()

#Log Transformation
qvc_data %>% 
  select(Package_Cnt) %>% 
  mutate(log_Package_Cnt = log(Package_Cnt + 1)) %>% 
  gather() %>% 
  ggplot(mapping = aes(x = value, group = key)) +
    geom_histogram(bins = 10) + 
    facet_wrap(~ key, scales = "free")
```

### Package_Scan_Dttm
```{r}
qvc_data %>% 
  select(Package_Scan_Dttm) %>% 
  skim()
```

### Carrier_Used_Tracking_Id
```{r}
qvc_data %>% 
  select(Carrier_Used_Tracking_Id) %>%
  mutate(Carrier_Used_Tracking_Id = as.character(Carrier_Used_Tracking_Id)) %>% 
  skim()

qvc_data %>% 
  select(Carrier_Used_Tracking_Id) %>%
  mutate(Carrier_Used_Tracking_Id = as.character(Carrier_Used_Tracking_Id)) %>% 
  summary()
```

### Product_Id
```{r}
qvc_data %>% 
  select(Product_Id) %>%
  mutate(Product_Id = as.character(Product_Id)) %>% 
  skim()

qvc_data %>% 
  select(Product_Id) %>%
  mutate(Product_Id = as.character(Product_Id)) %>% 
  summary()

qvc_data %>% 
  select(Product_Id) %>%
  mutate(Product_Id = as.character(Product_Id)) %>% 
  head()
```

```{r}
qvc_data %>% 
  mutate(Sales_Order_Nbr = as.character(Sales_Order_Nbr)) %>% 
  select(Sales_Order_Nbr) %>% 
  group_by(Sales_Order_Nbr) %>% 
  tally() %>% 
  arrange(desc(n))
```

