---
title: "Data Processing"
author: "Evan Canfield"
date: "7/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Library Calls
```{r library, warning=FALSE, message=FALSE}
if(!require(pacman)){install.packages("pacman")}
library(pacman)

p_load(  
  caTools,
  Hmisc, 
  corrplot,
  geosphere,
  leaflet,
  lubridate,
  noncensus,
  plotly,
  readxl,
  RColorBrewer,
  skimr,
  stringr,
  tidyverse
)
```

# Data Import
```{r cache=TRUE}
#Processed QVC Data Frame
qvc_data <- readRDS(file = "./data/output/Group Project Dataset/QVC_Data_Compiled.rds")

qvc_distctr <- read.csv(file = "./data/QVCdist_ctr.csv", stringsAsFactors = FALSE)

#US Population Census Data - 2018 PRojection
us_census_2018 <- read.csv("./data/us_census_population_change_2018_reformatted.csv")
```

## Check Data Import
### QVC Data
```{r}
#glimpse(qvc_data)
```

### Census Data
```{r}
#glimpse(qvc_distctr)
```

### Census Data
```{r}
glimpse(qvc_data)
```

# Drop Unnecessary Variables
```{r}
qvc_data <-
  select(
    Sales_Order_Nbr, 
    Order_Dt, 
    Total_Line_Amt,
    Source_Ship_Warehouse_Nbr,
    Delivery_Confirmation_Dt,
    SHIP_TO_ZIP
  )
```


```{r}
qvc_data %>% 
  group
```


# Subset Data For Quicker Process Times
```{r}
set.seed(101)

split_ratio <- 0.30

sample <- sample.split(Y = qvc_data$Sales_Order_Nbr
                       ,SplitRatio =  split_ratio)

qvc_data_sample = subset(qvc_data, sample == TRUE)
```

# Join Distribution Center Information
```{r}
qvc_data_1 <- qvc_data_sample %>% 
  left_join(
    select(qvc_distctr,Source_Ship_Warehouse_Nbr,POSTL_CD),
            by = "Source_Ship_Warehouse_Nbr"
            ) %>% 
  rename(zip_distctr = POSTL_CD)

#glimpse(qvc_data_1)
```


# Add Geographic Data
Latitude, Longitude, City, State will be added to the data frame in relation to the shipping warehouse and destination zip codes. FIPs codes will be added based on the detination zip codes as well.

## State Data Frame
First, a data frame listing each state along with the state's abbreviation, region, and division, based on R internal data, will be developed.
```{r}
# Develop data frame of state name, abbreviations, regions, and dvisions
state_df <- data.frame(State = state.name
                       , Abb = state.abb
                       , Region = state.region
                       )


#Re-define Regions from 4 to 5

Southwest <- c("Arizona", "New Mexico", "Texas", "Oklahoma")

state_df <- state_df %>% 
  mutate(Region = as.character(Region), 
         Region = if_else(Region == "North Central", "Midwest", Region),
         Region = if_else(Region == "South", "Southeast", Region), 
         Region = if_else(State %in% Southwest, "Southwest", Region))
```

## Join Location Data - Uncensus Package
```{r}
# Load Zip Code Data
data("zip_codes")

# Reinstate leading zeros for FIPs codes
zip_codes$fips  <- str_pad(string = zip_codes$fips, 
                                width = 5, 
                                side = "left", 
                                pad = "0")

# Join Zip Code Data
qvc_data_2 <- qvc_data_1 %>%
  
  #Join to Distribution Center Zip Code
  left_join(
    select(zip_codes, zip:longitude),
    by = c("zip_distctr" = "zip")
  ) %>% 
  
  #Rename Distribution Center Columns
  rename(
    city_distctr = city,
    state_distctr = state,
    lat_distctr = latitude,
    lon_distctr = longitude
  ) %>% 
  
  #Join to Desitnation Zip Code
  left_join(zip_codes
            , by = c("SHIP_TO_ZIP" = "zip")) %>% 
  
  #Rename Destination Columns
  rename(
    city_dest = city,
    state_dest = state,
    lat_dest = latitude,
    lon_dest = longitude,
    fips_dest = fips,
    zip_dest = SHIP_TO_ZIP
  ) 

#Reorder Columns
qvc_data_3 <- qvc_data_2 %>% 
  select(Sales_Order_Nbr,Package_Id,Party_Id:Product_Id,Ordered_Qty, Shipped_Qty, Actual_Total_Package_Qty, Merchandise_Dept_Desc,
         Order_Dt, Shipped_Dt, Rescheduled_Delivery_Dt, Delivery_Confirmation_Dt,
         Source_Ship_Warehouse_Nbr, city_distctr:lon_distctr, zip_distctr,
         city_dest:lon_dest, zip_dest, fips_dest
         )
glimpse(qvc_data_3)
```

# FUll Join Test
```{r}
# Load Zip Code Data
data("zip_codes")

# Reinstate leading zeros for FIPs codes
zip_codes$fips  <- str_pad(string = zip_codes$fips, 
                                width = 5, 
                                side = "left", 
                                pad = "0")

# Join Zip Code Data
qvc_data_2 <- qvc_data_1 %>%
  
  #Join to Distribution Center Zip Code
  left_join(
    select(zip_codes, zip:longitude),
    by = c("zip_distctr" = "zip")
  ) %>% 
  
  #Rename Distribution Center Columns
  rename(
    city_distctr = city,
    state_distctr = state,
    lat_distctr = latitude,
    lon_distctr = longitude
  ) #%>% 
  
glimpse(qvc_data_2)

qvc_data_fj <- qvc_data_2 %>% 
  #Join to Desitnation Zip Code
  full_join(zip_codes
            , by = c("SHIP_TO_ZIP" = "zip")) %>% 
  
  #Rename Destination Columns
  rename(
    city_dest = city,
    state_dest = state,
    lat_dest = latitude,
    lon_dest = longitude,
    fips_dest = fips,
    zip_dest = SHIP_TO_ZIP
  ) %>% 
  select(-SHIP_TO_STATE)

glimpse(qvc_data_fj)

qvc_data_fj %>% 
  distinct(fips_dest)

#Reorder Columns
qvc_data_fj_reorder <- qvc_data_fj %>% 
  select(Sales_Order_Nbr,Package_Id,Party_Id:Product_Id,Ordered_Qty, Shipped_Qty, Actual_Total_Package_Qty, Merchandise_Dept_Desc,
         Order_Dt, Shipped_Dt, Rescheduled_Delivery_Dt, Delivery_Confirmation_Dt,
         Source_Ship_Warehouse_Nbr, city_distctr:lon_distctr, zip_distctr,
         city_dest:lon_dest, zip_dest, fips_dest
         )
glimpse(qvc_data_fj_reorder)
```

```{r}
qvc_data_fj_reorder %>% 
  distinct(fips_dest) %>% 
  nrow()

zip_codes

zip_codes %>% 
  distinct(fips) %>% 
  nrow()  

qvc_data_fj_reorder_1 <- qvc_data_fj_reorder %>% 
  mutate(Total_Line_Amt = if_else(is.na(Total_Line_Amt),1, Total_Line_Amt))

summary(qvc_data_fj_reorder_1)
```


# New Variables

Several new variables will need to be developed for our analysis. The following code creates the following variables

* Fulfillment_Time: The number of days from order to delivery
* Rescheduled: Binary which indicates if an order has been rescheduled (1) or not (0)
* Distance: Distance, in miles, from Distribution Center to Desitination, based on longitude and latitude of the respective zip codes

## Fulfillment Time
```{r cache.lazy=TRUE}
qvc_data_4 <- qvc_data_3 %>% 
  mutate(Fulfillment_Days = as.numeric(Delivery_Confirmation_Dt - Order_Dt),           # Number of Days Order to Delivered
        Rescheduled = if_else(is.na(Rescheduled_Delivery_Dt), 0, 1) ,                  # Rescheduled = 1 (Binary)
        ) %>% 
  rowwise() %>% 
  mutate(Distance = round((distm(x = c(lon_distctr, lat_distctr)                       # Calculate Distance Between Distribution Center an Destination, Miles
                          , y = c(lon_dest, lat_dest)
                          , fun = distHaversine) * 0.000621371), 0)
  ) %>% 
  ungroup() %>%                                                                        # Remove Rowwise Property
  select(-Rescheduled_Delivery_Dt)                                                     # Drop Rescheduled Date Once Binary is Created

glimpse(qvc_data_4)
```


# Visualizations
## Checking Delivery Times 
```{r}
qvc_data_4 %>% 
  select(Sales_Order_Nbr,Order_Dt, Shipped_Dt,Delivery_Confirmation_Dt, Fulfillment_Days) %>% 
  filter(Fulfillment_Days < 1) %>% 
  nrow()
```

The 373 observations where the delivery date preceded the order date are a very small portion of the data and will be treated as an error. These observations are dropped

```{r}
qvc_data_5 <- qvc_data_4 %>% 
  filter(Fulfillment_Days >= 1) %>% 
  drop_na()
```

```{r}
p_Fulfillment_Days_hist <- ggplot(data = qvc_data_5,
                                     mapping = aes(x = Fulfillment_Days))+
  geom_histogram() +
  theme(legend.position = "none")

p_Fulfillment_Days_hist
```

Log Tranformed
```{r}
p_Fulfillment_Days_hist_log <- ggplot(data = qvc_data_5,
                                     mapping = aes(x = log(Fulfillment_Days)))+
  geom_histogram() +
  theme(legend.position = "none")

p_Fulfillment_Days_hist_log
```


# Project Question 1
### Sales vs Population
```{r}
# Develop data frame with total sales, average shippin gdistance, and average shipping time (days) for every state
qvc_sales_by_state <- qvc_data_5 %>% 
  group_by(state_dest) %>% 
  summarise(Total_Sales = round(sum(Total_Line_Amt), 0),
            Distance_Avg = round(mean(Distance),0),
            Fulfillment_Days_Avg = round(mean(Fulfillment_Days),1))

#Convert State to Character
qvc_sales_by_state$state_dest <- as.character(qvc_sales_by_state$state_dest)

# Join US Census population data frame with state dataframe for joining purposes 
state_pop <- us_census_2018 %>% 
  left_join(state_df, by = c("State" = "State")) %>% 
  select(State, Abb, Region, Division, Population) %>% 
  mutate(Population_per_k = Population / 1000)

#Convert factors to characters
state_pop$Abb <- as.character(state_pop$Abb)

state_pop$Region <- as.character(state_pop$Region)

state_pop$Region <- as.character(state_pop$Region)

state_pop$Division <- as.character(state_pop$Division)

# Join census frame with sales dataframe
qvc_sales_by_state_1 <- qvc_sales_by_state %>% 
  
  #Join US Census Data
  left_join(state_pop, by = c("state_dest" = "Abb")) %>% 
  
  #Rename Column
  rename(Abb = state_dest) %>% 
  
  #Create New Variables
  mutate(Sales_Percentage = round((Total_Sales / sum(Total_Sales)*100), 2),
         Sales_Per_Capita = round((Total_Sales / Population_per_k), 3)) %>% 
  
  #Arrange Data Frame Columns
  select(State, Abb, Region, Division, Population, Population_per_k, Total_Sales, Sales_Percentage, Sales_Per_Capita, Fulfillment_Days_Avg, Distance_Avg) %>% 

glimpse(qvc_sales_by_state_1)
```

## Visualize Sales vs Population
```{r, warning=FALSE}
p_sales_vs_pop <- ggplot(data = qvc_sales_by_state_1
       , mapping = aes(x = Population_per_k, y = Total_Sales)) +
  geom_point(aes(color = Region, text = State), size = 4, alpha = 0.9) +
  #geom_text(aes(label=if_else(Sales_Percentage > 4,as.character(State),'')),hjust=0,vjust=0) +
  geom_smooth(method = "lm", se = TRUE) +
  #xlim(0, 5e7)  +
  scale_colour_brewer(palette = "Set2") +
  labs(title = "Total Sales vs Population, Per State") +
  theme(legend.position = "top") + 
  theme_minimal()

ggplotly(p_sales_vs_pop
         , tooltip = "text") %>% 
   layout(legend = list(orientation = "h", x = 0.5, y = -0.2))

```

## Linear regression
```{r}
qvc_sales_by_state.lm <- lm(formula = Total_Sales ~ Population_per_k
                            , data = qvc_sales_by_state_1)

summary(qvc_sales_by_state.lm)

state_sales_rate_per_k <- unname(qvc_sales_by_state.lm$coefficients[2])
state_sales_rate_per_k
```

With the linear regression slope we create a dataframe documenting each state's delta from the mean, in sales
* Calcualte projected sales, and then calcualte actual - projected. Generate plot of each state, with postitive delta on one side of the midline, negative on the other


## Visualize Distance vs Delivery Times
```{r, warning=FALSE}
p_distance_vs_delivery_times_line <- ggplot(data = qvc_sales_by_state_1
       , mapping = aes(x = Fulfillment_Days_Avg, y = Distance_Avg)) +
  geom_point(aes(color = Region, text = State), alpha = 0.9) +
  geom_smooth(method = "lm", se = TRUE) +
  #xlim(0, 5e7)  +
  scale_colour_brewer(palette = "Set2") +
  labs(title = "Shipping Distance vs Fulfillment Length (Days), Per State") +
  theme(legend.position = "bottom") + 
  theme_minimal()

ggplotly(p_distance_vs_delivery_times_line
         , tooltip = "text") %>% 
  layout(legend = list(orientation = "h", x = 0.5, y = -0.2))
```

## Visualize Spend per Population Vs Delivery Times
```{r, warning=FALSE}
p_sale_per_pop_vs_delivery_times <- ggplot(data = qvc_sales_by_state_1
       , mapping = aes(x = Fulfillment_Days_Avg, y = (Total_Sales / Population_per_k))) +
  geom_point(aes(color = Region, text = State), alpha = 0.9) +
  geom_smooth(method = "lm", se = TRUE) +
  #xlim(0, 5e7)  +
  scale_colour_brewer(palette = "Set2") +
  labs(title = "Sales per Capita vs Fulfillment Length (Days), per State") +
  theme(legend.position = "bottom") +
  theme_minimal()

ggplotly(p_sale_per_pop_vs_delivery_times
         , tooltip = "text") %>% 
  layout(legend = list(orientation = "h", x = 0.5, y = -0.2))
```

## Vertical Line State PLot
### Develop Data Frame
```{r}
glimpse(qvc_sales_by_state_1)

state_sales_rate_per_k

qvc_sales_vert_line <- qvc_sales_by_state_1 %>% 
  select(State, Region, Population_per_k, Total_Sales) %>% 
  mutate(Projected_Sales = round(Population_per_k * state_sales_rate_per_k, 0),
         Sales_Delta = Projected_Sales - Total_Sales, 
         Sales_Direction = if_else(Sales_Delta > 0 , "Positive", "Negative"))

qvc_sales_vert_line

p0_sales_vert <- ggplot(
  data = qvc_sales_vert_line
  , mapping = aes(x = Sales_Delta
                  , y = reorder(State, Sales_Delta)
                  , color = Sales_Direction
                  )
) 

p1_sales_vert <- p0_sales_vert + 
  geom_vline(xintercept = 0, color = "gray30") +
  geom_point() + 
  theme_minimal()

p1_sales_vert
  
p2_sales_vert <- p1_sales_vert +
  facet_wrap(~ Region, ncol = 2, scales="free_y") +
    guides(color=FALSE) + labs(x = "Sales Delta", y = "") +
    theme(axis.text=element_text())

p2_sales_vert

```


## Plot Distribution Centers
```{r}
qvc_data_5

# Distribution Center Location Dataframe
distrctr_location <- qvc_data_5 %>% 
  select(Source_Ship_Warehouse_Nbr:zip_distctr) %>% 
  distinct() %>% 
  arrange(state_distctr, city_distctr)

# Distribution Center Sales Data
distrctr_sales <- qvc_data %>% 
  group_by(Source_Ship_Warehouse_Nbr) %>% 
  summarise(Total_Sales = round(sum(Total_Line_Amt), 0)) %>% 
  mutate(Total_Sales_Percent = round((Total_Sales / sum(Total_Sales) * 100), 1)) %>% 
  left_join(distrctr_location, by = "Source_Ship_Warehouse_Nbr") %>% 
  arrange(desc(Total_Sales))

```

Inspect the new data frames
```{r}
glimpse(distrctr_location)

glimpse(distrctr_sales)
```

Leaflet Visual
```{r}
leaflet(data = distrctr_sales) %>% 
  addTiles()%>% 
  addCircleMarkers(lng = ~Longitude_distctr
                   , lat = ~Latitude_distctr
                   #, stroke = FALSE
                   , radius = ~Total_Sales_Percent
                   , label = paste(distrctr_sales$City_distctr,", ",(distrctr_sales$State_distctr))
                   )%>% 
  setView(lng = -96, 
          lat = 37.8, 
          zoom = 4) %>%
  setMaxBounds(lng1 = -180, 
               lng2 = -60,
               lat1 = 73, 
               lat2 = 15)
```

# Export
```{r}

glimpse(qvc_data_5)
state_df

# Full Data Frame
write.csv(x = qvc_data_5
          , file = "./Tableau Visuals/data/qvc_data_complete.csv")

#State Data Frame
write.csv(x = state_df
          , file = "./Tableau Visuals/data/state_df.csv")

# Vertical Line Plot
write.csv(x = qvc_sales_vert_line
          , file = "./Tableau Visuals/data/qvc_sales_vert_line.csv")
```

