---
title: "Question 1 - Point Path Map"
author: "Evan Canfield"
date: "7/24/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

# Library 
```{r library}
if(!require(pacman)){install.packages("pacman")}
library(pacman)

p_load(  
  caTools,
  Hmisc, 
  geosphere,
  lubridate,
  noncensus,
  sf,
  skimr,
  stringr,
  tidycensus,
  tidyverse,
  tigris
)
```

# Import Data Frame
```{r }
qvc_path <- readRDS(file = "./data/output/q1_tableau/qvc_data_q1.rds")
```

```{r}
glimpse(qvc_path)
summary(qvc_path)
```

# Check if Party ID is associated with a single location
```{r}
qvc_path%>%
  select(Party_Id, zip_dest) %>% 
  group_by(Party_Id) %>%
  mutate(n = n_unique(zip_dest)) %>% 
  arrange(desc(n)) %>% 
  distinct()
```

The Party_Id / Zip Code connection is not unique. Therefore Party_Id / Warehouse Number cannot be used as a unique path. Zip_dest must be used as the unique destination identifier 
# Drop Unnecessary Variables
```{r}
qvc_path_1 <- qvc_path %>% 
  select(
      - fips_dest
    , -Fulfillment_Days
    , -Distance
  ) %>% 
  mutate(path_code = paste(Party_Id, zip_dest,Source_Ship_Warehouse_Nbr, sep = "_"),
         destination_code = paste(Party_Id, zip_dest, sep = "_")
         ) %>% 
  rename(origin_code = Source_Ship_Warehouse_Nbr) %>% 
  select(Sales_Order_Nbr:Merchandise_Dept_Desc,
         origin_code:destination_code,
         -zip_distctr
         )

glimpse(qvc_path_1)
```

```{r}
qvc_path_2 <- qvc_path_1 %>% 
    rename(
      city_origin = city_distctr,
      state_origin = state_distctr,
      lat_origin = lat_distctr,
      lon_origin = lon_distctr
    )
  
glimpse(qvc_path_2)  
```

# Create a Unique Row Identifier
```{r}
qvc_path_3 <- qvc_path_2 %>% 
  mutate(id = rownames(qvc_path_2))

skim(qvc_path_3)
```


# Break Data Frame into Origin and Destination Data Frame
```{r}
qvc_path_org_1 <- qvc_path_3

qvc_path_dest_1 <- qvc_path_3

skim(qvc_path_3)
```

# Origin Dataset
```{r}

glimpse(qvc_path_org_1)

qvc_path_org_2 <- qvc_path_org_1 %>% 
  select(
         id,
         Sales_Order_Nbr, 
         city_origin:lon_origin,
         path_code,
         origin_code,
         destination_code,
         Total_Line_Amt,
         state_dest, 
         zip_dest
         ) %>% 
  rename(
    city = city_origin,
    state = state_origin,
    lat = lat_origin,
    lon = lon_origin
  ) %>% 
  mutate(origin_destination = "origin")

glimpse(qvc_path_org_2)
```

# Destination Dataset
```{r}
qvc_path_dest_2 <- qvc_path_dest_1 %>% 
  select(
         id,
         Sales_Order_Nbr, 
         city_dest:lon_dest,
         path_code,
         origin_code,
         destination_code,
         Total_Line_Amt
         ) %>% 
    rename(
    city = city_dest,
    state = state_dest,
    lat = lat_dest,
    lon = lon_dest
  ) %>% 
  mutate(state_dest = state,
    origin_destination = "destination")

# Reorder to Match
qvc_path_dest_3 <- qvc_path_dest_2 %>% 
  select(id:state,
         lat:state_dest,
         zip_dest,
         origin_destination)

glimpse(qvc_path_dest_3)
```

# Bind Origin and Destination Data Frames
```{r}
qvc_path_comb <- rbind(qvc_path_dest_2, qvc_path_org_2)

qvc_path_comb <- qvc_path_comb %>% 
  arrange(id, desc(origin_destination))

# Reorder
qvc_path_comb <- qvc_path_comb %>% 
  select(id:destination_code  ,
         state_dest:origin_destination,
         Total_Line_Amt,
         -id)

glimpse(qvc_path_comb)
```

# State Centroid Lat/Lon
```{r}

zip_codes

qvc_path_comb

zip_codes_state_centroid <- zip_codes %>% 
  group_by(state) %>% 
  summarise(lat_center = mean(latitude),
            lon_center = mean(longitude))
```

# Add Centroid Lat/Lon Data
```{r}
zip_codes_state_centroid

qvc_path_comb_1 <- qvc_path_comb %>% 
  left_join(zip_codes_state_centroid
            , by = "state")

glimpse(qvc_path_comb_1)

head(qvc_path_comb_1)
```

# Replace Lat / Lon For Destination With State Centroid
```{r}
qvc_path_state_centroid <- qvc_path_comb_1 %>% 
  mutate(lat = if_else(origin_destination == "destination", lat_center, lat),
         lon = if_else(origin_destination == "destination", lon_center, lon))

glimpse(qvc_path_state_centroid)
```


# Export
```{r}
write.csv(x = qvc_path_comb_1
          , "./data/output/q1_tableau/qvc_path_points_all.csv"
          , row.names = FALSE)

write.csv(x = qvc_path_state_centroid
          , "./data/output/q1_tableau/qvc_path_points_centroid.csv"
          , row.names = FALSE)
```

