---
title: "Question 3 - Exploratory"
author: "Evan Canfield"
date: "7/26/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Library Calls

The following libraries are used to develop the data frames for Question 1 of the Group Project

```{r library, warning=FALSE, message=FALSE}
if(!require(pacman)){install.packages("pacman")}
library(pacman)

p_load(  
  caTools,
  Hmisc, 
  geosphere,
  lubridate,
  randomForest,
  skimr,
  stringr,
  tidyverse
)
```

# Data Import
```{r cache=TRUE}
#Processed QVC Data Frame - Sample
qvc_data_sample <- readRDS(file = "./data/output/Group Project Dataset/qvc_data_compiled_new_variables_sample.rds")

#Processed QVC Data Frame
qvc_data_all <- readRDS(file = "./data/output/Group Project Dataset/qvc_data_compiled_new_variables.rds")
```


```{r}
glimpse(qvc_data_sample)
```


```{r}
glimpse(qvc_data_all)
```

# Create a Customer Based Data Set
```{r}
#Sample
qvc_customer_order_count_sample <- qvc_data_sample %>%
  distinct() %>% 
  select(Party_Id, Sales_Order_Nbr) %>% 
  group_by(Party_Id) %>%
  count() %>% 
  arrange(-n)

qvc_customer_order_count_sample
```

```{r}
#All

qvc_customer_order_count_all <- qvc_data_all %>%
  distinct() %>% 
  select(Party_Id, Sales_Order_Nbr) %>% 
  group_by(Party_Id) %>%
  count() %>% 
  arrange(-n)

qvc_customer_order_count_all
```

# Sample
```{r}
qvc_customer_repeat_sample <- qvc_customer_order_count_sample %>% 
  mutate(Repeat = if_else(n > 1, 1, 0)) %>% 
  select(-n)

glimpse(qvc_customer_repeat_sample)

summary(qvc_customer_repeat_sample)
```

# All
```{r}
qvc_customer_repeat_all <- qvc_customer_order_count_all %>% 
  mutate(Repeat = if_else(n > 1, 1, 0)) %>% 
  select(-n)

glimpse(qvc_customer_repeat_all)

summary(qvc_customer_repeat_all)
```

# Export Repeat Customer
## Sample
```{r}
saveRDS(object = qvc_customer_repeat_sample
        , file = "./data/output/q3_modeling/qvc_customer_repeat_sample.RDS")

write.csv(x = qvc_customer_repeat_sample
          , "./data/output/q3_modeling/qvc_customer_repeat_sample.csv",
          row.names = FALSE)
```

## Sample
```{r}
saveRDS(object = qvc_customer_repeat_all
        , file = "./data/output/q3_modeling/qvc_customer_repeat_all.RDS")

write.csv(x = qvc_customer_repeat_all
          , "./data/output/q3_modeling/qvc_customer_repeat_all.csv",
          row.names = FALSE)
```

# Import Model Data
```{r}
qvc_customer_model <- read.csv(file = "./data/output/q3_modeling/model_dataset_no_labels.csv"
                               , stringsAsFactors = FALSE)
```

# Inspect
```{r}
# summary(qvc_customer_model)
# 
glimpse(qvc_customer_model)
```

# Convert Data Types for Imported Data Frame
```{r}

qvc_customer_model_1 <- qvc_customer_model %>% 
  select(-X) %>% 
  mutate(Sales_Order_Nbr = as.character(Sales_Order_Nbr),
         Package_Id = as.character(Package_Id),
         Party_Id = as.character(Party_Id),
         Product_Id = as.character(Product_Id),
         Source_Ship_Warehouse_Nbr = as.character(Source_Ship_Warehouse_Nbr),
         Order_Dt = as.Date(Order_Dt),
         Shipped_Dt = as.Date(Shipped_Dt),
         Rescheduled_Delivery_Dt = as.Date(Rescheduled_Delivery_Dt),
         Delivery_Confirmation_Dt = as.Date(Delivery_Confirmation_Dt)
         )

glimpse(qvc_customer_model_1)
```

# Join Repeat With Model Data
```{r}

qvc_customer_model_repeat <- qvc_customer_model_1  %>% 
  left_join(qvc_customer_repeat
            , by = "Party_Id")

glimpse(qvc_customer_model_repeat)
```

# Random Forest
# Convert Character to Factor
```{r}
qvc_customer_model_repeat_fct <- qvc_customer_model_repeat %>% 
  mutate_if(is.character, as.factor)
  
```

#Drop Variables
Character / Factor variables with more than 53 categories will be dropped. These are generally unique ids
```{r}
glimpse(qvc_customer_model_repeat_fct)

qvc_customer_model_repeat_fct_trim <- qvc_customer_model_repeat_fct %>% 
  select(-Sales_Order_Nbr,
         -Package_Id,
         -Party_Id,
         -Product_Id,
         -city_distctr,
         -city_dest,
         -zip_distctr,
         -zip_dest,
         -Source_Ship_Warehouse_Nbr
         )

glimpse(qvc_customer_model_repeat_fct_trim)
```

# Export Random Forest Data Frame
```{r}
saveRDS(object = qvc_customer_model_repeat_fct_trim
        , "./data/output/q3_modeling/qvc_data_model_rf.RDS")
```


## Subset
```{r}
set.seed(101) 

sample = sample.split(qvc_customer_model_repeat_fct_trim$Repeat, SplitRatio = .70)

df_forest.train = subset(qvc_customer_model_repeat_fct_trim, sample == TRUE)

df_forest.valid = subset(qvc_customer_model_repeat_fct_trim, sample == FALSE)
```

```{r}
rf <- randomForest(Repeat ~ .
                   , data = df_forest.train
                   , importance = TRUE
                   , ntree = 20)
```


