---
title: "Question 3 - Exploratory"
author: "Evan Canfield"
date: "7/26/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Library Calls

The following libraries are used to develop the data frames for Question 1 of the Group Project

```{r library, warning=FALSE, message=FALSE}
if(!require(pacman)){install.packages("pacman")}
library(pacman)

p_load(  
  caTools,
  caret,
  corrplot, 
  Hmisc, 
  geosphere,
  lubridate,
  randomForest,
  ROCR,
  skimr,
  stringr,
  tidyverse
)
```

# Data Import
```{r cache=TRUE}
#Processed QVC Data Frame - Sample
qvc_data <- read.csv(file = "./data/output/q3_modeling/model_dataset_with_labels.csv"
                            , stringsAsFactors = FALSE)

qvc_data_complete <- readRDS(file = "./data/output/Group Project Dataset/qvc_data_compiled_new_variables.rds")
```

```{r}
glimpse(qvc_data_complete)
```


```{r}
qvc_data <- qvc_data %>% 
  select(-X)
```


# Determine Fulfillment Days Mean / Median
```{r}
median(qvc_data$Fulfillment_Days)

mean(qvc_data$Fulfillment_Days)

ggplot(data = qvc_data
       , mapping = aes(log(Fulfillment_Days, 2))) +
  geom_density()
```

 
# Logistic Regression- Drop Variables
# And Filter AK, HI, PR
```{r}

state_list <- c("PR", "AK", "HI")

qvc_data_1 <- qvc_data %>% 
  filter(!state_dest %in% state_list) %>% 
  select(Repeat,
         Total_Line_Amt,
         Unit_Price_Amt, 
         Ordered_Qty,
         Shipped_Qty,
         Actual_Total_Package_Qty,
         Merchandise_Dept_Desc,
         Rescheduled:Distance
         )
```

# Develop Early / Late Binary
```{r}
qvc_data_2 <- qvc_data_1 %>% 
  mutate(Late = if_else(Fulfillment_Days >  7, 1, 0))
```


# Data Type Conversion
```{r}
qvc_data_3 <- qvc_data_2 %>% 
  mutate(
    Merchandise_Dept_Desc = as.factor(Merchandise_Dept_Desc),
    Repeat = as.factor(Repeat),
    Rescheduled = as.factor(Rescheduled)
  )
```

```{r}
glimpse(qvc_data_3)
```

# Histograms
```{r}
 qvc_data_3 %>%
   keep(is.numeric) %>% 
   gather() %>% 
  ggplot(aes(value)) +
     facet_wrap(~ key, scales = "free") +
     geom_histogram(bins = 50)
```

# Correlation Matrx
```{r}
 qvc_data_3_num <- qvc_data_3 %>%
   keep(is.numeric)

qvc_cm <- cor(qvc_data_3_num)

corrplot(qvc_cm)

```

# Processing
Drop Ordered / Shipped, Days, Unit Price, Log Transform Total Line Amount
```{r}
qvc_data_4 <- qvc_data_3 %>% 
  select(  -Unit_Price_Amt
         , -Ordered_Qty
         , -Shipped_Qty
         )
```


# Subset
```{r}
# Data partition
set.seed(101) 

sample = sample.split(qvc_data_4$Repeat, SplitRatio = .70)
df.train = subset(qvc_data_4, sample == TRUE)
df.valid = subset(qvc_data_4, sample == FALSE)

```


# Baseline
```{r}
model.lr_baseline <- glm(formula = Repeat ~ .
    , binomial(link='logit')
    , data = df.train)

summary(model.lr_baseline)

exp(model.lr_baseline$coefficients)
```


```{r}
#Prediction Output - Baseline Logistic Regression
predict_reg <- predict(model.lr_baseline 
                          , newdata = df.valid 
                          , type='response')

#Convert to Dataframe
predict_reg <- data.frame(Probability = predict_reg, row.names = NULL)

#Create Predicition Binary Based on Threshold
threshold <- 0.5

predict_reg <- predict_reg %>% 
  mutate(Results = if_else(Probability >= threshold, 1, 0)) %>%
  mutate(Results = as.factor(Results)) %>% 
  mutate(Known_Valid = df.valid$Repeat) %>% 
  mutate(Model = "Logistic Regression")

#Confusion Matrix
confusionMatrix(data = predict_reg$Results
                , reference =  df.valid$Repeat)
```

#### ROC Curve
```{r}
pred_reg <- prediction(predict_reg$Probability, df.valid$Repeat)

perf_reg <- performance(pred_reg, "tpr", "fpr")
plot(perf_reg)
```

#### AUC
```{r}
# Calculate and print AUC value
auc_reg <- performance(pred_reg, measure="auc")
auc_reg <- auc_reg@y.values[[1]]
print(paste("AUC for the regression model is:",  round(auc_reg,3)))
```